---
title: "Analysis of short cyclic peptides"
author: "Sebastian Worms"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    highlight: tango  # specifies the syntax highlighting style
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = '', fig.width = 6, fig.height = 6)

library(tidyverse)
library(knitr)
library(ggplot2)
library(Peptides)
library(writexl)
library(gridExtra) # To align multiple plots
library(readxl) # Read excel files

#Load some custom functions
source("../functions/function_analysis.R")

```

# Introduction

This is the analysis of the peptides from the drift experiment, looking the the induced, NNK8 file, the only one that showed correlation between synonymous sequence. The required datasets were created in the "peptide_analysis.R" file. The encoding of enrichment ratio is 'NA' in cases where no reads were present in the first generation and some in the fifth (giving an 'infinite' enrichment ratio).

```{r 'load datasets', echo=FALSE}
directory <- "C:/Users/worms/ngs_data/2022_06_07_drift_seq/90-666155004b/00_fastq/all_files"

# Read a sample of the set
count_set <- read.csv2(file = file.path(directory, "count_set_params.csv"))  %>%
  filter(library == "nnk" & condition == "induced") %>% # Take only NNK, induced peptides
  mutate(enrichment_ratio = ifelse(is.infinite(enrichment_ratio),NA,enrichment_ratio)) %>% # Going to drop all the infinite enrichment ratio to NA
  filter(grepl("^TGC([ATGC][ATGC][GT]){7}$", seq)) # Drop the ones with sequences that don't fit
```

## Sequence quality and reads distribution

```{r 'sequence_quality'}
unique_seq <- count_set %>%
  pull(seq) %>%
  unique() %>% 
  length()

total_count <- count_set %>% select(gen1,gen5) %>% sum()

unique_peptide <- count_set %>%
  pull(standard_seq) %>%
  unique() %>% 
  length()

```

The data set contains ``r unique_seq` different DNA sequences, sequenced with `r total_count` reads. As expected the size of the library and sequencing means most sequences are sequenced a few times only, with few sequences vastly over represented (hundreds, or in very rare case over a thousand reads), either due to selection directly or potentially quality issues in the randomization of the primers. 

```{r}
table_gen1 <- count_set %>% select(gen1) %>% summary() %>% data.frame()
table_gen5 <- count_set %>% select(gen5) %>% summary() %>% data.frame()
kable(table_gen1 , caption = "Distribution of reads in the first generation")
```

After five days of drift, the distribution is very skewed: most sequences aren't detected at all and have been purged. A small fraction of the sequences have been highly enriched with counts in the thousands of even hundreds of thousands. 

A cursory glance shows that those sequences that were enriched have high counts even at gen1, compatible with the selection theory of initial imbalance in ratios. Let's have a closer look at that.

### Are common sequences in gen1 more likely to be selected?

I would like to investigate whether the sequences with a higher counts in the first generation tends have a higher ratio. One issue to be taken is sequences that aren't present in both generation of the drift as by disappearing they could skew my ratios. 

```{r }
plot_init <- ggplot(count_set %>% filter(gen1 > 0 & gen5 > 0), aes(y = gen1, x = enrichment_ratio_log)) +
  geom_point(alpha = 0.05)

plot_hist <- ggplot(count_set, aes(x = enrichment_ratio_log)) +
  geom_histogram()

plot_init %>% 
  print %>%
  suppressWarnings()


plot_hist %>% 
  print %>%
  suppressWarnings()
```

If anything, the opposite is observed: sequences with a higher count in gen 1 have a lower enrichment ratio. The line appearance is likely due to the ratios you observe with 1 or 2 reads in gen5. That's slightly artefactual though. What's interesting is that a lot of the more highly enriched sequence also have gen1 counts. So that's compatible with the theory that they're pre-selected, even if most common gen1 sequences are common for random reasons and have negative fitness.

The weird spikes are due to the pseudocounts.

## Homonyms sequence

```{r 'Adding a number of seq'}

# count_set <- count_set %>%
#   group_by(standard_seq) %>%
#   mutate(homonym_seq = n()) %>%
#   mutate(average_enrich_ratio = psych::geometric.mean(enrichment_ratio)) %>%
#   ungroup()

multiple_seq <- count_set %>%
  filter(homonym_seq > 1) %>%
  pull(standard_seq) %>%
  unique() %>%
  length()
```



One good sign that a signal is real is if it's observed in multiple sequence. We compute a 'standard sequence' from the peptide sequence to account for cyclization. The very large size of potential octamers means only a relatively small numbers of peptides are encoded by multiple genes = `r multiple_seq` out of `r unique_peptide`.

We can look at those, but we're limiting ourselves to a small subset of peptides.

## Homonyms with stop codons

One prediction is that peptides with stop codons should show lower correlation between synonymous sequence as the peptide isn't expressed. Let's see.

```{r 'Synonymous peptides with stop codons'}
ratio_list <- count_set %>%
  filter(homonym_seq > 1) %>%
  get_enrichment_list()

ratio_list_stop <- count_set %>%
  filter(homonym_seq > 1) %>%
  filter(!is.infinite(enrichment_ratio_log)) %>%
  filter(grepl("\\*", standard_seq)) %>%
  get_enrichment_list()

ratio_list_no_stop <- count_set %>%
  filter(homonym_seq > 1) %>%
  filter(!is.infinite(enrichment_ratio_log)) %>%
  filter(!grepl("\\*", standard_seq)) %>%
  get_enrichment_list()

cor(ratio_list$X1, ratio_list$X2)
cor(ratio_list_stop$X1, ratio_list_stop$X2)
cor(ratio_list_no_stop$X1, ratio_list_no_stop$X2)

df_cor_ratios <- data.frame(has_stops = logical(),
                            X1 = double(),
                            X2 = double())


ratio_list_stop <- ratio_list_stop %>%
  data.frame() %>%
  mutate(has_stops = TRUE)

ratio_list_no_stop <- ratio_list_no_stop %>%
  data.frame() %>%
  mutate(has_stops = FALSE)

df_cor_ratios <- rbind(df_cor_ratios, ratio_list_stop, ratio_list_no_stop)

plot_corr_stops <- ggplot(data = df_cor_ratios, aes(x = X1, y = X2)) +
  geom_point(alpha = 0.05) +
  labs(title = "Correlation between the enrichment ratios of sequences encoding for the same peptide sequence") +
  xlab("Enrichment ratio 1") +
  ylab("Enrichment ratio 2") +
  geom_smooth(method = lm, se = FALSE) +
  facet_grid(~ has_stops)

print(plot_corr_stops) %>% suppressWarnings()
```

Correlation is still pretty positive. Could it be that the issues is that when there some of the very low reads are just misread of common ones (or mutation?)

## Let's look at the top sequences with confirmation

```{r 'Ordering sequence'}
count_set_multiple <- count_set %>%
  filter(homonym_seq >= 3) %>%
  arrange(desc(standard_seq))

count_set_multiple %>% arrange(desc(average_enrich_ratio)) %>% head(n = 100) %>% select(1,2,3,6,7,10,13)


#Going to check that the pseudocount don't interfere with correlation
cor_count_set_no_peud <- get_enrichment_list(count_set %>%
                                               filter(gen1 > 0.5 & gen5 > 0.5))
# Correlation is actually better than when we don't take in the counts with zeroes.
cor(cor_count_set_no_peud$X1,cor_count_set_no_peud$X2)

plot_cor <- ggplot(data= cor_count_set_no_peud, aes(x=X1, y=X2)) +
  geom_point(alpha= 0.05) +
  geom_smooth(method = "lm", se = FALSE)

plot_cor %>% print()

#Generally decent!
```

The correlation with pseudocounts is quite good! Not perfect of course, but you have a real, repeated, correlated decrease or increase.


```{r 'Adding logP and charge'}
# #Using the peptide package to add estimates of hydrophobicity and charge and some other parameters
# count_set <- count_set %>%
#   mutate(hydrophob = hydrophobicity(standard_seq)) %>%
#   mutate(charge = charge(standard_seq)) %>%
#   mutate(pi = pI(standard_seq)) %>%
#   mutate(aindex = aIndex(standard_seq)) %>%
#   mutate(boman = boman(standard_seq))
# 
# # write.csv2(count_set, file= file.path(directory,"count_set_params.csv"))
# count_set <- read.csv2(file.path(directory,"count_set_params.csv"), row.names = FALSE)
```



```{r}
peptide_candidate <- count_set %>%
  filter(!grepl('\\*', standard_seq)) %>%
  filter(homonym_seq > 2) %>%
  pull(standard_seq) %>%
  unique() %>%
  length()

count_set_candidate <- count_set %>%
  filter(!grepl('\\*', standard_seq)) %>%
  filter(homonym_seq > 2) %>%
  arrange(standard_seq) %>%
  arrange(desc(average_enrich_ratio))


# write_xlsx(count_set_candidate, file.path(directory,"candidate_peptides.xlsx"))
```

## Charge, pI, aliphatic index, hydrophobicity

None of those characteristics seems to show a clear correlation with enrichment.


```{r 'Looking at hydrophobicity'}
count_set_sample <- count_set[sample(1:nrow(count_set), 100000, replace=FALSE),]

p1 <- ggplot(data = count_set_sample, aes(x = hydrophob, y = enrichment_ratio_log)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm", se = TRUE)

p2 <-ggplot(data = count_set_sample, aes(x = pi, y = enrichment_ratio_log)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm", se = TRUE)

p3 <-ggplot(data = count_set_sample, aes(x = charge, y = enrichment_ratio_log)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm", se = TRUE)

p4 <-ggplot(data = count_set_sample, aes(x = aindex, y = enrichment_ratio_log)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm", se = TRUE)

p5 <-ggplot(data = count_set_sample, aes(x = boman, y = enrichment_ratio_log)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm", se = TRUE)

grid.arrange(p1, p2, p3, p4, p5, ncol = 2)
```

## Presence of RR

While they're both purged, we seem to have a slightly lower enrichment ratio for peptides without RR motifs, although with a lot of overlap. Should be noted that this only holds with one single RR, not two. 

```{r}
count_set <- count_set %>%
  mutate(RR = grepl(pattern = "RR", x = standard_seq))

count_set <- count_set %>%
  mutate(Rcount =  str_count(standard_seq, "RR")) %>%
  mutate(Rcount = as.factor(Rcount))

p7 <- ggplot(data = count_set, aes(x = Rcount, y = enrichment_ratio_log)) +
  geom_violin() +
  xlab("Peptide containing RR motifs") +
  ylab("Log2 of the enrichment ratio")

p7

stat_test <- t.test(x = count_set %>% filter(RR == TRUE) %>% pull(enrichment_ratio_log),
       y = count_set %>% filter(RR == FALSE) %>% pull(enrichment_ratio_log))

anova_test <- aov(enrichment_ratio_log ~ Rcount, data = count_set)

shapiro.test(count_set[1:5000,]$enrichment_ratio_log)
bartlett.test(enrichment_ratio_log ~ Rcount, data = count_set)

mean_values2 <- aggregate(enrichment_ratio_log ~ Rcount, data = count_set, geometric.mean)

```
```{r}
count_set %>% select(c(Rcount,enrichment_ratio_log)) %>% group_by(Rcount) %>% lapply(. %>% pull(enrichment_ratio_log),psych::geometric.mean())
```

