---
title: "Competition simulations"
output: html_document
date: "2023-08-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r 'loading functions', include= FALSE}
folder_loc <- getwd()
source(file = file.path(folder_loc,"simulation_functions.R"))
```


# Introduction
This report aims to compare the observed ratio of peptides in the competitions experiment to what we know of their mumaxes and see if they are compatible.

## Input parameters
The models requires a bunch of parameters, both strain-specific and generic. I'll input the generic parameters now. K and µ0 will be obtained from the data of the experiment of 2023.06.17.

```{r 'generic parameters', include = TRUE}
# Importing the data
params_df <- read.csv(file.choose())

# Take the average of all K for the carrying capacity
K <- params_df %>% filter(induction == "ara") %>% pull("K") %>% mean()
print(paste0("The value of K is ", round(K, digits = 2)))

# Same with the h0 metabolic rate parameter
h0 <- params_df %>% filter(induction == "ara") %>% pull("h0") %>% mean()
print(paste0("The value of h0 is ", round(h0, digits = 2)))
```
The model also require a df with peptides names, mumaxes and y0

```{r 'input df'}
input_df <- data.frame(peptide = params_df %>% 
                         arrange(peptide) %>% 
                         pull(peptide) %>% 
                         unique(),
                       mumax =  params_df %>% 
                         filter(induction == "ara") %>% 
                         group_by(peptide) %>%
                         summarise(mumax_mean = mean(mumax)) %>%
                         ungroup() %>% pull(mumax_mean),
                       y0 = params_df %>%
                         pull(y0) %>%
                         .[1]
)

print(input_df)
```
## Running the model

Once we have the parameters we can run the model.Here for exemple is a representation of a 16 hours competition.

```{r 'running_model'}
hours = 16
time_sim <- seq(0,hours*3600, by = 100)

result_sim <- competition_simulation(input_df, time = time_sim, K = K, h0 = h0)
```

```{r 'Graphing the simulation'}
data_graph <- pivot_simulation_df(result_sim) %>%
  mutate(time = time /3600) # convert time to hours

plot_exemple <- ggplot(data_graph, aes(x= time, y = OD, color = peptide)) +
  geom_point(alpha = 0.5, na.rm = TRUE) + 
  xlab("Time (hours)") +
  ylab("Optical density at 600 nm") +
  scale_x_continuous(breaks = c(0,4,8,12,16)) +
  ggtitle("Simulated growth of the peptides over 4 hours.")

plot_exemple
```

